Lattice Boltzmann simulations of magnetically responsive\\ anisotropic particle stabilized emulsions

### 2025-03-04 chapters/paper2.tex
Changes 1. Changed chapter 2 content 2. Addeed chapter 1 curvature results

Previous work demonstrated that applying electric fields to silica-stabilized emulsion droplets caused particle stabilizers to unjam and then rejam, locking the emulsion in a new morphology. \cite{cui_stabilizing_2013} Considering the potential applications of an already synthesized porous material with tunable microstructure, the second part of this work will explore using magnetic fields to manipulate bijels after synthesis. \cite{vanoli_bijels_2022, cha_bicontinuous_2019} Bijels simulated under no field had a constant magnetic in the z direction of strength $\Bar{B} = 0, 0.2, 0.5, 1$ with their microstructure and particle order analyzed.

Underlying mechanism being rearrangement of particles on the interface in response to magnetic field, or unjamming and rejamming characterized by a change in the number of bonded neighbors. Explains behaviour seen in differences in the system above and below the isotropic to nematic transition?

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/results/paper2/domain_size.png}
    \caption{Left plot shows the average domain size $(L_1)$ in markers and the nematic order parameter $(S)$ in lines versus time. Right plot shows the same data rescaled over $t_{iso \rightarrow nematic}$.}
    \label{fig:P2_domain_scaling}
\end{figure}

In this system, the mechanism of microstructure modification after formation is expected to be due to magnetic field driven particle ordering changes. Figure \ref{fig:P2_domain_scaling} demonstrates that the average domain size $(L_1)$ does not change significantly with the applied field, with a slight increase of about $\approx 0.3 L/R$, noted to be within error. However, clues into the underlying mechanics are visible when examining the nematic order parameter lines overlaid on the domain size markers. At $\Bar{B} = 0.2$, the domain size change corresponds with the nematic order parameter change, indicating a direct link between domain size change and particle ordering. At $\Bar{B} > 0.2$, a gap appears between the domain size change and the nematic order parameter change, suggesting steric effects influence the rate of bijel microstructure change. The timescales of these rearrangements may also impact the observed microstructure changes, which can be characterized by the domain size change over time rescaled by a characteristic timescale.

A naive guess as to the dominating timescale of the response would be the isotropic to nematic time of the system or when $S \geq 0.3$. The right plot of Figure \ref{fig:P2_domain_scaling} demonstrates the time evolution of the system when rescaling time with the isotropic to nematic transition time, $t_{iso \rightarrow nematic}$. From this plot, it can be seen that application of the field causes collapse of domain size changes at different field strengths on top of one another, showing that this naive guess is a correct one. The changes in domain size at with no applied field can be attributed to the reordering of particles at the interface, leading to coarsening at long time scales consistent with what Gunther et al. observed. \cite{gunther_timescales_2014}

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/results/paper2/domain_size_aniso.png}
    \caption{Plots of the perpendicular $(L_{\perp})$ and parallel $(L_{\parallel})$ domain sizes over time rescaled over $t_{iso \rightarrow nematic}$}
    \label{fig:P2_domain_aniso}
\end{figure}

In the previous section, even if the average domain size did not change much, there was domain size anisotropy found upon application of the field. This observation and trends seen remains true when applying the field after formation as can be observed in \ref{fig:P2_domain_aniso}. Time was again rescaled with the isotropic to nematic transition time for each system, demonstrating that the timescale of response is controlled using this timescale, and is independent from field strength. The domain size change observed appears to be about $1 L/R$ less than what can be found when applying the same field size during bijel synthesis as the increase in domain size for $L_{parallel}$ for prolate particles is $\approx 2 L/R$ which is $50 \%$ lower than what was found. This indicates that microstructure control after synthesis is not as effective. This might be due to steric effects being much stronger after synthesis, limiting how much particles can move or tilt before rejamming. Despite this, the domain size change found is well within what has been seen to be effective in the literature, indicating that magnetic field enabled in-situ microstructure modifications can be utilized within bijels. \cite{cha_bicontinuous_2019, khan_nanostructured_2022, vanoli_bijels_2022}

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/results/paper2/interface_angle.png}
    \caption{Plots of the average angle between the particle dipole moment and interface normal $(\psi)$ over time. While there are minor changes in for the oblate particles, these are not as significant as the ordering of particles to the field.}
    \label{fig:P2_interface_angle}
\end{figure}

To determine if the interface follows the orientation of the particles, the average angle between the particle symmetry axis and the interface normal $(\psi)$ was plotted against time rescaled by $t_{iso \rightarrow nematic}$ in Figure \ref{fig:P2_interface_angle}. From ths plot, it can be observed that upon application of the magnetic field, the particle tilts out of the interface at an equal normalized rate. However, the angle reached before the particles begin returning to their equilibrium configuration is magnetic field strength dependent, suggesting that the degree of microstructure changed observed might be dependent on how much the particle tilts away from its equilibrium position. At non-equilibrium particle tilts, there is more leeway for coarsening to occur before rejamming occurs, resulting in larger microstructural differences.

Future work will more closely examine the role of the particles in observed microstructure changes by calculating the average angle between the particle symmetry axis and the interface normal. In addition, the impact of initial order of the particle monolayer on the degree of microstructure change observed will be characterized. This will be accomplished by using bijels that have been simulated under magnetic fields and changing the applied field. Example conditions would be a bijel being simulated under a field strength of $\Bar{B} = 0.2$ and increasing the applied field to $\Bar{B} = 1$ or switching off the applied field. Based on the findings so far it is predicted that initial particle order and microstructure change are inversely correlated. Additionally, the timescales of these microstructure changes will be investigated. When switching a field on, there are magnetic field, steric and interfacial forces that govern the dynamics. However, when reducing the field strength, the magnetic field derived force isn't present, resulting in steric and interfacial forces playing a significant role. These differences can be examined by investigating the timescales the observed response occurs at, as well as the particle properties over time.

% \textcolor{blue}{Still needs additional explanation of how using templates made under fields and decreasing and increasing field strengths will change results}

>>>>

\section{Results}\label{sec:results_p2}
\subsection{Field strength dependence on domain
size}\label{field-strength-dependence-on-domain-size}

Bijels have proposed applications as soft matter based material
templates, drug delivery vectors and separations systems. The materials
performance in each application is dependent upon the microstructure of
the bijel template. Many of the applications listed above benefit from
stimuli response, allowing for in-situ microstructure tuning to modify
permeability of the material. We want to understand whether bijels
provide access to reversible structural modifications which would allow
stimuli responsive bijels to be used in in-situ applications.
Additionally, in many of the proposed stimuli responsive applications
for bijels, the timescales of response is important to characterize and
would also provide a mechanistic understanding of the reasons why the
microstructure changes occur. We probe these aspects of structural
response of bijels in this section by assessing the response of bijels
stabilized by particles with volume fraction of \(\phi_p = 0.1\)
simulated under no applied field to an applied field in the range
\(\bar{B} = 0, 0.2, 0.5, 1\). We begin our investigation by visualizing
the structure before and after the application of magnetic fields.

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/microstructure_viz-field_on.png} 
\caption{Visualizations of bijels stabilized by oblate and prolate particles simulated under no fields at $t = 0$ (left columns) and $t = 10^5$ (right columns). The first and third row detail the microstructure evolution for oblate and prolate particle stabilized bijels with no applied field while the second and fourth rows show bijels stabilized by oblate and prolate particles respectively with a field strength of $\bar{B} = 1$ applied. Particle reorientation to the direction of the field can be seen upon application of the magnetic field, resulting in microstructure changes driven through domain coarsening until the particles jam. We see similar behavior in bijels stabilized by oblate particles upon application of the magnetic field.}
\label{fig:microstructure_viz-field_on} 
\end{figure}

In Figure \ref{fig:microstructure_viz-field_on} we show representative
snapshots of the bijel at the initial and final timestep with
\(\bar{B}_z = 0\) and \(\bar{B}_z = 1\). We show that upon application
of the field, particles align to the direction of the field. As the
particles orient to the field, they disrupt the packing of particles at
the interface, causing domain coarsening as the particles rearrange into
new positions. The particles then rejam in place once the interfacial
area and particle cross sectional area match. Domain coarsening observed
in the figure can be explained through the ordering of particles at the
interface, causing a reduction in the available area that the particles
have for stabilization.

We characterize the microstructure changes observed through calculating
a characteristic length scale, called the average domain size. The
average domain size is calculated using the first moment of the 3D
averaged structure factor. Past work has shown that this method
correctly reproduces the scaling expected in phase separating systems.
The structure factor is calculated from the fourier transform of the
fluctuations in the order parameter. \cite{kendon_inertial_2001}

\begin{equation}
S(\mathbf{k}, t) = \widetilde{\phi}(\mathbf{k}, t)\widetilde{\phi}(\mathbf{-k}, t)
\end{equation}

\(\mathbf{k}\) represents the wave vector in each direction of each
lattice grid point. The fluctuations in the order parameter are
calculated from the order parameter of the bijel defined as
\(\widetilde{\phi} = \phi - \langle \phi \rangle\). \(\phi\) is
calculated through subtraction of the red from the blue fluid density,
\(\phi = \rho^b - \rho^r\). The average domain size is commonly
calculated using the spherically averaged structure factor where all
points corresponding to \(k = |\mathbf{k}|\) are binned, where \(k\) is
the magnitude of \(\mathbf{k}\).

\begin{equation}
S(k, t) = \frac{1}{n} \sum_{k - \Delta < k < k + \Delta} S(\mathbf{k}, t)
\label{eq:structure_factor_calc}
\end{equation}

The average domain size can then be calculated from the spherically
averaged structure factor after calculating the first moment of the
structure factor using the following expression.

\begin{equation}
L(t) = 2\pi \frac{\sum S(k, t) }{ \sum k S(k, t) }
\end{equation}

We found that this method for determining domain size generally
overestimated the observed channel diameter. However, previous studies
have demonstrated that the scaling behavior of the first-moment-derived
domain size aligns with the expected phase separation in spinodal
decomposition. The overestimation arises due to the limited number of
samples at small k values, where the bin size is roughly equivalent to
the k value. Additionally, in simulations with periodic domains, both
the structure factor and computed length scale are influenced by
finite-size effects. Nonetheless, this measurement still provides
valuable insight into trends in the average domain size.

To correlate the microstructure evolution to the ordering of the
particle monolayer to the magnetic field, we calculate the nematic order
parameter. \cite{cuesta_monte_1990} First, the average orientational
tensor is calculated from the particle orientations in Eq
\ref{eq:orientational_order_tensor}

\begin{equation}
Q = \frac{1}{n_p} \sum_{i} (\frac{3}{2}[\hat{\mathbf{o}}_i \otimes \hat{\mathbf{o}}_i] - \frac{1}{2}\mathbf{I})
\label{eq:orientational_order_tensor}
\end{equation}

In Eq \ref{eq:orientational_order_tensor}, \(\hat{\mathbf{o}}_i\)
defines the orientation vector of particle \(i\) and \(\mathbf{I}\) is
the identity matrix. The nematic director or average particle
orientation is the eigenvector corresponding to the largest eigenvalue
of \(Q\) which is the value of the nematic order parameter, \(S\). The
assembly of particles can be said to have nematic order if
\(S \geq 0.3\) and is fully aligned to the director if \(S = 1.0\).
\cite{cuesta_monte_1990}

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/domain_size-field_on.png} 
\caption{Plotting the spherically averaged domain size normalized with $R_p$($L/R$) of the particle at different magnetic field strengths, $\bar{B}$, and the nematic order parameter $S$ in lines. We quantify the domain size and ordering of the particles, and show that the change in both parameters are correlated with the strength of the applied magnetic field.} 
\label{fig:domain_size-field_on} 
\end{figure}

In Figure \ref{fig:domain_size-field_on}, we present the time evolution
of the normalized average domain size, where normalization is performed
using the volume-equivalent spherical particle radius \(R = 7.9\). The
results indicate that upon applying a magnetic field to a bijel
initially formed without one, the final average domain size increases by
up to 5.2\% and 2.5\% for oblate- and prolate-particle-stabilized bijels
respectively. Additionally, in the absence of an external field, domain
coarsening occurs for both particle morphologies. However, the results
for oblate particles with a magnetic field is attributed to coalescence
of domains rather than magnetic field driven ordering of the particles.

In the same figure, the time evolution of the nematic order parameter
\(S\) is also plotted to illustrate changes in particle alignment.
Initially, at \(\bar{B}_z = 0\), a gradual increase in \(S\) is
observed, attributed to steric interactions driving particle reordering.
This effect contributes to the domain coarsening seen in the absence of
an applied field, consistent with the findings of Günther et
al.~\cite{gunther_timescales_2014}. In the case of
oblate-particle-stabilized bijels without an external field, domain
coalescence is primarily driven by steric interactions rather than
field-induced rearrangement. Furthermore, the rate and extent of
particle alignment to the field exhibit a dependence on field strength,
as evidenced by the earlier onset of nematic ordering with increasing
\(\bar{B}\). We also observed between the time evolution of the nematic
order parameter and domain size, particularly when comparing results for
\(\bar{B}_z = 0.5, 1\) against those for \(\bar{B}_z = 0.2\). This
suggests that nematic ordering alone may not fully account for the
observed changes in domain size.

Previous studies have demonstrated that microstructural properties,
particularly length scale and tortuosity, play an important role in the
performance of porous materials across various applications, including
battery electrodes and tissue engineering. \cite{zhang_promoting_2019}
\cite{ebner_tortuosity_2014} \cite{hsieh_architected_2021} Specifically,
lower tortuosity and more ordered ion-conducting domains have been shown
to enhance charge/discharge rates in battery systems, while controlled
tortuosity in artificial bone grafts facilitates improved cell growth
and nutrient transport.

Precise characterization of the anisotropic microstructure is essential
for the tailored design of application-specific porous materials. To
quantify the characteristic length scales in each Cartesian direction,
we compute the second-moment-derived structure factor without radial
averaging. \cite{jansen_bijels_2011} Following the structure factor
calculation described in Eq. \ref{eq:structure_factor_calc}, the second
moment of the structure factor in direction \(\beta\) is determined for
each Cartesian coordinate axis as

\begin{equation}
L_{\beta}(t) = 2 \pi \sqrt{\frac{\sum_{\mathbf{k_{\beta}}}  S(\mathbf{k}, t) }{\sum_{\mathbf{k_{\beta}}}  k_{\beta}^2S(\mathbf{k}, t)}}
\label{eq:second_moment_domain}
\end{equation}

From Eq \ref{eq:second_moment_domain}, we can calculate the length scale
in each cartesian direction, \(L_x, L_y, L_z\). Here, we make a further
refinement simplifying the domain sizes to directions parallel and
orthogonal to the magnetic field. These result in the length scales and
their associated calculation, \(L_{\parallel} = L_{z}\) and
\(L_{\perp} = \frac{L_x + L_y}{2}\). Tortuosity has been defined using
the geometric, diffusional and hydraulic tortuosities. Here we define
the diffusional tortuosity, \(\tau = \frac{D}{D_{eff}}\) which
characterizes the ratio between the free diffusion of a tracer and the
effective diffusion of a tracer in the material.

To calculate the tortuosities of the structures generated from our
simulations, we first calculate \(\phi\) of the system over time from
the filled densities. Many bijels are synthesized from monomer diffused
into the oil phase, or the oil phase itself is intrinsically
polymerizable. Our next step is to binarize the order parameter with the
condition, \(\phi \leq 0 = 0\) and \(\phi > 0 = 1\) which reflects a
solid red fluid phase and void blue fluid. We then mask all unconnected
components, using the \texttt{label} function in \texttt{skimage}
leaving only the percolating domain representing the bijel.
\cite{van2014scikit} We utilize the python package taufactor which
solves the diffusion equation for the structure, calculating the time
taken for a tracer to go from one side of the system to the other.
\cite{kench_taufactor_2023} We utilize the periodic boundary condition
solver for this investigation. In previous work, we identified that the
length scale derived from the second moment of the structure factor is
only useful for identifying length scale anisotropy. Therefore, we plot
the directional second moment derived length scales and tortuosity in
Figure \ref{fig:domain_size_aniso-field_on}.

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/domain_size_aniso-field_on.png} 
\caption{Plotting the anisotropic domain size and tortuosities against the applied magnetic field strength on the left and right respectively for bijels stabilized with magnetically responsive oblate and prolate ellipsoids. We see that upon application of the field, bijels stabilized with prolate particles experience a domains ize increase in the direction parallel to the applied field direction, $L_{\parallel}$, while the domain size perpendicular to the applied field, $L_{\perp}$ decreases. The opposite is true for bijels stabilized with oblate particles. The tortuosity data demonstrate that as the domain size increases, the tortuosity decreases and vice versa.} 
\label{fig:domain_size_aniso-field_on} 
\end{figure}

From Figure \ref{fig:domain_size_aniso-field_on}, we see that domain
size anisotropy is observed upon application of the field for both
particle morphologies. We also see some domain anisotropy with no
applied field, which increases upon application of the magnetic field.
The directional domain sizes, \(L_{\perp}\) and \(L_{\parallel}\)
increase by \(73\%\) and \(7\%\) respectively for oblate particles and
\(10\%\) and \(44\%\) respectively for oblate particles. The axes that
coarsens the most is consistent with past work, explained through the
directions particles reorient in response to the applied magnetic field.
As particles unjam and rejam at the interface due to reorientating to
the magnetic field, they adopt direction specific surface coverages
which manifest as anistropic domains upon jamming. In Figure
\ref{fig:domain_size_aniso-field_on}, we demonstrate that the
application of the magnetic field changes the domain size. The
tortuosity of the bijel is initially around \(\tau \approx 1.5\),
consistent with simulations of the diffusive tortuosities of gyroid
structures using Lattice Boltzmann simulations.
\cite{luo_macroscopic_2020} Upon application of the field,
\(\tau_{\perp}\) slightly decreases while \(\tau_{\parallel}\)
increases. We have seen that an increasing domain size decreases the
tortuosity in our past work. \cite{karthikeyan_formation_2024} This
observation is in line with the domain size changes observed in Figure
\ref{fig:domain_size_aniso-field_on}. The results show that the
tortuosity of the bijel is noticeably affected by the application of
magnetic fields. This data suggests that magnetic fields applied onto
bijels stabilized with ellipsoidal particles undergo domain anisotropy
upon application of magnetic fields. However the particle morphology,
specifically its axis of symmetry, affect the direction that the domain
anisotropy is characterized.

Next, we turn to investigate the time gap between the average domain
size increase and the nematic order parameter. We hypothesized that the
time taken for particle reorientation to the field would control the
degree and rate of domain size change when the bijel begins with
particles with random order. This result suggests that the dynamics of
the system at higher magnetic field strengths are dependent upon other
factors that vary upon application of the magnetic field. To understand
this phenomena, we focus on the dynamics of the particle monolayer.
First, we visualize the particle monolayer of the bijels stabilized with
prolate ellipsoids at the initial and final configuration before and
after application of a magnetic field.

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/particle_viz-field_on.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with no applied field. The second and fourth rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. Particle reorientation to the direction of the field can be seen, resulting in coarsening of the domains.} 
\label{fig:particle_viz-field_on} 
\end{figure}

In Figure \ref{fig:particle_viz-field_on}, we see the effect of the
application of the magnetic field on the monolayer. The initial
configuration of particles at the monolayer show that the orientations
are random and that the particles tend to lie flat on the interface.
However, upon application of the magnetic field the particles orient to
the field direction. The effect that Gunther. et al.~observed for bijels
stabilized with anisotropic particles can also be observed in these
snapshots as the particles can be observed to have more local order in
the final state than in the initial state with no applied field. We
observe that for bijels stabilized by prolate ellipsoids, we see the
adoption of more regular arrangements of the colloids on the interface.
We investigate the effect of local particle ordering next. Past work by
Bresme and Faraudo and Davies investigating the reorientation of
particles at interfaces under magnetic fields have shown that a critical
magnetic field exists, above which the particle orients to the field and
below which the particle orientation is a function of the capillary
force and magnetic field derived force. \cite{davies_interface_2014}
\cite{bresme_orientational_2007} Particle adsorption into the interface
with microparticles have also been shown to be irreversible due to the
high adsorption energy. Upon particle reorientation, the microstructure
change can be driven solely through particle unjamming and re-jamming,
or it can be driven through the particles dragging the interface during
this process. The mechanism of unjamming and re-jamming can be
investigated using the average interfacial angle, \(\psi\).

The interface angle of each particle is defined as
\(\psi = \arccos{\frac{ \hat{\mathbf{o_i}}\cdot \hat{\mathbf{n_i}}}{|\hat{\mathbf{o_i}}| |\hat{\mathbf{n_i}}|}}\)
where \(\hat{\mathbf{n_i}}\) is the normal of the interface closest to
the center of the particle \(i\) and \(\hat{\mathbf{o_i}}\) is the
orientation of particle \(i\) which follows the symmetry axis of the
ellipsoid. For prolate particles, they prefer to lie flat on the
interface, meaning that with no field \(\psi = 90^{\circ}\).
\(\hat{\mathbf{n_i}}\) is calculated by first using the particle filling
algorithm described and calculating the order parameter. Next, a mesh of
the isocontour at \(\phi = 0\) is generated using the marching cubes
algorithm in scikit-image with a box length of 4. \cite{van2014scikit}
Next, the vertex closest to the center of particle \(i\) was found.
\(\hat{\mathbf{n_i}}\) was then obtained from the normal corresponding
to the vertex identified in the previous step.

Another factor affected by the application of the magnetic field is the
Steinhardt 6 fold order parameter, \(Q_6\), which characterized the
local particle ordering. \cite{steinhardt_bond-orientational_1983} It
has been used to characterize crystallization, glass transitions and
crystal structures for colloidal and molecular crystals. We utilize the
average across all particles on the interface, \(\langle Q_6 \rangle\)
and track this parameter as a function of time. Past work has identified
that local crystallization of colloidal crystals takes place at
\(\langle Q_6 \rangle = 0.38\), and that
\(\langle Q_6 \rangle^{HCP} = 0.485\).
\cite{steinhardt_bond-orientational_1983} \cite{toxvaerd_role_2020}
\cite{mickel_shortcomings_2013} We utilize a voronoi cell derived
definition of the neighbor list of each particle, addressing some of the
shortcomings highlighted in the original distance based definition.
\cite{steinhardt_bond-orientational_1983}
\cite{mickel_shortcomings_2013} We first define a complex orientational
vector, \(q_{lm}\)

\begin{equation}
q_{lm}(i) = \frac{1}{N_b(i)} \Sigma_{j = 1}^{N_b(i)} Y_{lm}(\vec{R}_{ij})
\end{equation}

Where \(N_b\) is the number of nearest neighbors of particle \(i\),
\(l\) is controls to the degree of spherical harmonic to be calculated,
\(m\) is an integer defined as \(-l \leq m \leq l\) that defines which
spherical harmonic is being calculated and \(\vec{r}_{ij}\) is the
distance between two particle centers. The neighbor list \(N_b\) is
usually defined by a cutoff distance from the center of particle \(i\)
to its nearest neighbors \(j\). Steinhardt then defined a bond order
parameter \(q_l\) from \(q_{lm}\) as

\begin{equation}
Q_{l}(i) = \sqrt{\frac{4 \pi}{2l + 1} \Sigma_{m = -l}^{l} |q_{lm}(i)|^2}
\end{equation}

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/interface_angle-nint-field_on.png} 
\caption{Plots of the interface angle $(\langle \psi \rangle)$ and six fold Steinhardt order parameter $(\langle Q_6 \rangle)$ on the left and right respectively, versus time for bijels stabilized with oblate and prolate ellipsoids on the top and bottom rows respectively. We demonstrate that upon application of a magnetic field, the behavior of $(\langle \psi \rangle)$ and $(\langle Q_6 \rangle)$ differ significantly between oblate and prolate ellipsoids indicating that the dynamics for both particle morphologies differ. We also observe magnetic field dependence in each property for both particle types, suggesting that these changes are driven primarily through the change in the applied field strength.} 
\label{fig:interface_angle-nint-field_on} 
\end{figure}

From Figure \ref{fig:interface_angle-nint-field_on}, we identify that
prolate and oblate particles respond differently at the interface. For
bijels stabilized by oblate ellipsoids, we characterize an increase in
\(\langle \psi \rangle\) from an initial value of \(19 ^{\circ}\),
before it reduces slightly and begins to increase once again. The
initial starting value of \(\langle \psi \rangle\) indicates the
particles are lying close to flat on the interface, before they begin to
tilt out of the interface when the magnetic field is applied. Then,
\(\langle \psi \rangle\) plateaus as the interface begins to move in
response to the tilting particle at the interface. This is caused by the
large adsorption energy of the particles at the interface of the bijel,
caused by high surface tension. Finally, the particles jam in place,
locking the interface at an angle to the particle. We see the rate of
increase in \(\langle \psi \rangle\) at early stages and the time spent
in the plateau increase with a larger applied magnetic field. The final
angle at the interface is also dependent upon the applied field
strength.

When analyzing \(\langle \psi \rangle\) for bijels stabilized by prolate
particles, this behavior can also be observed. \(\langle \psi \rangle\)
begins at \(80 ^{circ}\). The difference in the initial
\(\langle \psi \rangle\) compared to oblate particles is due to the
direction of the axis of symmetry. In prolate particles, the axis of
symmetry lies parallel with the interface and \(\langle \psi \rangle\)
measures the angle between the interface normal and the orientation of
the axis of symmetry. We see \(\langle \psi \rangle\) decrease slightly,
before plateauing and slowly increasing again. The rate of decrease in
\(\langle \psi \rangle\) is field dependent, although given the small
decrease in \(\langle \psi \rangle\) it is unclear if the final value of
this parameter is field dependent for bijels stabilized with prolate
particles. The changes in \(\langle \psi \rangle\) for bijels stabilized
with prolate particles are small compared to bijels stabilized with
oblate particles. We believe this is due to local particle effects at
the interface. We probe this by identifying the local particle ordering
at the interface characterized by \(\langle Q_6 \rangle\).

We see that for oblate particles, \(\langle Q_6 \rangle\) sees a
decrease from its initial value. However the dynamics are controlled by
the applied magnetic field strength. We split the phases of dynamics
into three sections, similar to \(\langle \psi \rangle\) with the first
section seeing a decrease in \(\langle Q_6 \rangle\) upon application of
the magnetic field, followed by a plateau and finally a gradual increase
before jamming takes place. These results indicate that the ordering of
the particles is reduced by the applied magnetic field, before a
restoration of some of the pre-existing order. The final values of
\(\langle Q_6 \rangle\) are magnetic field dependent, with a stronger
magnetic field having a lower final \(\langle Q_6 \rangle\) value
indicating that either the ordering of particles influence the angle at
the interface or vice versa. We observe that for bijels stabilized by
oblate particles, the interface ordering of the particles increases as a
function of the applied field strength. We also observe from plotting
\(\langle Q_6 \rangle\) that application of a magnetic field creates a
local crystallization of the colloid monolayer characterized as
\(\langle Q_6 \rangle > 0.38\). \cite{toxvaerd_role_2020}

Figure \ref{fig:domain_size-field_on}, we observed a gap between the
growth of the nematic order parameter and an increase in the domain
size. To probe why this occurs, we recast \(\langle \psi \rangle\),
\(\langle Q_6 \rangle\) and \(S\) in terms of the physical phenomena
they represent. The nematic order parameter represents the ordering of
particles to an average direction, which in the case of the application
of a magnetic field, is the field direction. Therefore, a timescale
reliant on the nematic order parameter would represent a timescale of
response rate limited to the response to the magnetic field, which
indicates a particle-field orientation dependence. We name this
timescale \(\tau_S\). When applying a field to a bijel system with a
randomly oriented particle monolayer, the most significant transition
point is the isotropic to nematic transition or when \(S \geq 0.3\). We
define \(\tau_S\) for this section of the analysis by identifying when
\(S \geq 0.3\).

We can also do the same for \(\langle \psi \rangle\) and
\(\langle Q_6 \rangle\). A timescale derived from
\(\langle \psi \rangle\) named \(\tau_{\langle \psi \rangle}\)
represents the capillary timescale between the interface and the
particle. Given the shape of the plot, we assign the
\(\tau_{\langle \psi \rangle}\) to be the time when the first minimum of
\(\langle \psi \rangle\) is observed for prolate particles and the first
maximum for oblate particles. A timescale derived from
\(\langle Q_6 \rangle\), named \(\tau_{\langle Q_6 \rangle}\) is also
calculated which represents a timescale of local particle rearrangement.
We define \(\tau_{\langle Q_6 \rangle}\) as the time when
\(\langle Q_6 \rangle \geq 0.38\) for prolate particles and the global
minimum of \(\tau_{\langle Q_6 \rangle}\) for the oblate particles. We
omit \(\bar{B} = 0\) as previous results have identified the domain size
changes observed there. We then plot the average domain size \(L\)
against time rescaled with these three timescales in Figure
\ref{fig:domain_size-field_on-scaled}.

\begin{figure} 
\centering 
\includegraphics[scale=0.35]{../figures/results/paper2/domain_size-field_on-scaled.png} 
\caption{Plots of domain size with different timescales. The top and bottom row represent the average domain size for bijels stabilized with oblate and prolate particles respectively. The figures on the left plots the domain size against time rescaled with the nematic to isotropic transition time, $\tau_S$. The figures in the middle plot the domain size against time rescaled with the interfacial capillary timescale, $\tau_{\langle \psi \rangle}$. The figures on the right plot the domain size against time rescaled with the particle rearrangement timescale, $\tau_{\langle Q_6 \rangle}$. We show that the controlling timescale when applying a magnetic field onto bijels with disordered particles is the latter timescale, indicating that the dominating effect is steric hindrance of the particles during rearrangment at the interface.} 
\label{fig:domain_size-field_on-scaled} 
\end{figure}

In Figure \ref{fig:domain_size-field_on-scaled}, the left plot plots the
average domain size against time rescaled using \(\tau_S\). We see that
the domain sizes do not collapse onto a single line, which is expected
if we want to explain the dynamics using the nematic order parameter.
This confirms the observation made above that the dynamics do not only
depend on the reorientation of particles to the field. The middle plot
demonstrates the average domain size against time rescaled using
\(\tau_{\langle \psi \rangle}\). We see that there is good agreement
between all simulations for prolate particles, although the point when
\(\frac{t}{\tau_{\langle \psi \rangle}} = 1\) lies when the domain size
transition has started. However for oblate particles we see that this
timescale does not describe the dynamics well.

The right plot demonstrates the domain size change after rescaling time
with \(\tau_{\langle Q_6 \rangle}\) where we see good collapse between
all plots while the domain size transition also begins at
\(\frac{t}{\tau_{\langle Q_6 \rangle}} = 1\). The latter point
demonstrates that the dynamics for the structural response of randomly
oriented particles are controlled by the local particle rearrangement at
the interface rather than the orientation of the particle to the
magnetic field.We see that before \(\tau_{\langle Q_6 \rangle} = 1\),
\(\bar{B} = 0.2\) for the oblate particles does not collapse onto the
other plots, indicating that the dynamics before that point are
controlled by another phenomena.

From the evidence in Figure \ref{fig:domain_size-field_on-scaled}, two
types of interfacial behaviors are characterized in addition to that
identified by Gunther et al.~\cite{gunther_timescales_2014} The first
type is what is observed for \(\bar{B} = 0.2\), where the interfacial
coverage and average orientation of the particle to the field and to the
interface change in lock step. This causes the interface to reorient at
the same rate as the particles reorienting to the magnetic field. The
second is observed for \(\bar{B} = 0.5, 1\), where there is a time lag
between particle ordering, and the intiation of domain coarsening. In
the latter scenario, at early times the particles are jolted out of
position before the large adsorption energy of the particles into the
interface causes the interface to move to the particle position, causing
domain coarsening as the interfacial area reduces.

\subsection{Increasing the applied
field}\label{increasing-the-applied-field}

We saw in the hysteresis curve that the microstructure change at higher
field strengths began to reduce with each subsequent field strength
jump, suggesting that the ordering of the particles were linked to the
domain size change observed. In the previous section, we saw that domain
size changes in bijels after jamming are constrained by magnetic field
driven local crystallization of the particle monolayer. In this section,
we explore how pre-existing order in the particles affect the degree of
domain size change observed. We apply fields of strength
\(\bar{B}_{template} = 0, 0.2, 0.5, 1\) during phase separation to
create bijels stabilized with ellipsoidal particles with pre-existing
particle order. We then apply a field strength of \(\bar{B}_z = 1\) and
observe the time evolution of the system for \(t = 10^5\) timesteps. We
simplify our nomenclature here demonstrated with an example.
\(\bar{B}_z: 0.0 \rightarrow 1.0\) represents a bijel made with no
field, and a field of \(\bar{B}_z = 1.0\) applied after the structure
has reached steady state. We begin by visualizing the microstructure
changes experienced by bijels simulated with pre-existing particle order
in Figure \ref{fig:microstructure_viz-field_up}.

\begin{figure}
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/microstructure_viz-field_up.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 0.2$. The second and fourth rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. We increase the applied field to $\bar{B}_z = 1$ in both cases and show that particle reorientation to the field occurs at the $\bar{B}_z = 0.2$ but not in the $\bar{B}_z = 1$ case.}
\label{fig:microstructure_viz-field_up}
\end{figure}

In Figure \ref{fig:microstructure_viz-field_up} we show representative
snapshots of a bijel stabilized with prolate particles at the initial
and final timestep with field conditions
\(\bar{B}_z: 0.2 \rightarrow 1.0\) and
\(\bar{B}_z: 1.0 \rightarrow 1.0\) in the top and bottom rows
respectively. We show that upon application of the field, the fluid
domains undergo a field strength dependent microstructrue response, with
the degree of response dependent upon the difference between the initial
and final magnetic field. We also observe a qualitative increase in the
alignment of the particles to the magnetic field for the
\(\bar{B}_z: 0.2 \rightarrow 1.0\) case. We quantify these observations
by plotting the average domain size and nematic order parameter in
Figure \ref{fig:domain_size-field_up}.

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/domain_size-field_up.png} 
\caption{Plot of the spherically averaged domain size normalized with $Rp(L/R)$ of the particle in markers and the nematic order parameter in lines. We show that the domain size change measured is correlated to the change in the particle ordering, characterized using the nematic order parameter.} 
\label{fig:domain_size-field_up} 
\end{figure}
